---
title: "Read and write in parallel"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Read and write in parallel}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
  knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
devtools::load_all()
```

By default, reads and writes are performed sequentially (i.e., not in parallel).
Users can opt-in to parallel read/write functionality via `options`.

## Simulate slow operations

```{r}
SlowDirectoryStore <- R6::R6Class("SlowDirectoryStore",
  inherit = DirectoryStore,
  public = list(
    get_item = function(key) {
      Sys.sleep(0.5) # Simulate a slow read.
      return(super$get_item(key))
    },
    set_item = function(key, value) {
      Sys.sleep(0.5) # Simulate a slow write.
      return(super$set_item(key, value))
    }
  )
)
```

## Read in parallel

Provide an integer >= 2 to the option to use forking-based parallelism.
This value will be passed to the `cl` parameter of `pbapply::pblapply`.

```{r}
options(pizzarr.parallel_read_enabled = 4)

root <- pizzarr_sample("dog.ome.zarr")
store <- SlowDirectoryStore$new(root)
zarr_arr <- zarr_open(store = store, path = "/0")
arr <- zarr_arr$get_item("...")$data
sum(arr)
```

## Write in parallel

```{r}
options(pizzarr.parallel_write_enabled = 4)

root <- pizzarr_sample("dog.ome.zarr")
store <- SlowDirectoryStore$new(root)
zarr_arr <- zarr_open(store = store, path = "/0")
arr <- zarr_arr$get_item("...")$data
zarr_arr$set_item("...", arr * 2.0)
doubled_arr <- zarr_arr$get_item("...")$data
sum(doubled_arr)
```

## Parallel operations with future backend

To use the `future` backend for `pbapply`, set the value of the option to the string `"future"`.

Cluster-based:

```{r}
options(pizzarr.parallel_read_enabled = "future")

cl <- parallel::makeCluster(2)
future::plan(future::cluster, workers = cl)

root <- pizzarr_sample("dog.ome.zarr")
store <- SlowDirectoryStore$new(root)
zarr_arr <- zarr_open(store = store, path = "/0")
arr <- zarr_arr$get_item("...")$data
sum(arr)

parallel::stopCluster(cl)
```

Multisession-based:

```{r}
options(pizzarr.parallel_read_enabled = "future")

future::plan(future::multisession, workers = 4)

root <- pizzarr_sample("dog.ome.zarr")
store <- SlowDirectoryStore$new(root)
zarr_arr <- zarr_open(store = store, path = "/0")
arr <- zarr_arr$get_item("...")$data
sum(arr)
```

## Sequential operations

To return to sequential mode, run:

```{r}
options(
    pizzarr.parallel_read_enabled = FALSE,
    pizzarr.parallel_write_enabled = FALSE
)
```

## Disable progress bar

Parallel operations are implemented with `pbapply`.
To disable the progress bar, run:

```{r}
pbapply::pboptions(type = "none")
```

To re-enable, run:

```{r}
pbapply::pboptions(type = "timer")
```